A New Model to Predict Survival in Patients with End-Stage Liver Disease

### 2022-07-03 analysis/article.Rmd
fix: cystatin c

  **Results**:
      Penalized regression algorithms yield the highest prediction performance in our ML algorithm benchmark.
      In favour of a simpler model we chose the least absolute shrinkage and selection operator (lasso) regression.
      Beside the classical MELD variables INR and bilirubin the lasso regression selected
      cystatin c over creatinine, and additionally IL-6, total protein and cholinesterase.
      The new model offers a better discrimination and calibration than MELD and MELD with sodium (MELD-Na) or the MELD-Plus7 risk score.

@@

The resulting model AMPEL model of end-stage liver disease
(AMELD) uses `r length(cfs)` predictors.
Beside the classical MELD variables INR and (direct) bilirubin the lasso
regression selected cystatin c over creatinine, and
additionally IL-6, total protein and cholinesterase.

@@

Our model, AMELD, uses `r length(cfs)` predictors.
Half of them, INR, (direct) bilirubin, and cystatin c are identical
or similar to the labortory measurements utilized in the classical MELD.
The direct bilirubin selected by our approach and the total bilirubin in MELD
are highly correlated (spearman correlation:
`r round(cor(amelddata$x, method = "spearman")["BILI_S", "BILID_S"])`) and
therefore interchangeable.

@@

In conclusion we developed a new model to predict the 90-day survival of
patients during the evaluation process for liver transplantation.
Our model, AMELD, extends the classical MELD predictors, INR,
(direct) bilirubin, cystatin c (instead of creatinine) by
synthesis parameters total protein and cholinesterase and
most importantly the inflammatory marker IL-6.
Using these `r length(cfs)` laboratory measurments the AMELD
outperforms the established MELD-Na and the recently published
MELD-Plus7 risk scores.
For a wider adoption a prospective multi center study for model
recalibration and validation is needed.

>>>>

  **Results**:
      Penalized regression algorithms yield the highest prediction performance in our ML algorithm benchmark.
      In favour of a simpler model we chose the least absolute shrinkage and selection operator (lasso) regression.
      Beside the classical MELD variables INR and bilirubin the lasso regression selected
      cystatin C over creatinine, and additionally IL-6, total protein and cholinesterase.
      The new model offers a better discrimination and calibration than MELD and MELD with sodium (MELD-Na) or the MELD-Plus7 risk score.

@@

The resulting model AMPEL model of end-stage liver disease
(AMELD) uses `r length(cfs)` predictors.
Beside the classical MELD variables INR and (direct) bilirubin the lasso
regression selected cystatin C over creatinine, and
additionally IL-6, total protein and cholinesterase.

@@

Our model, AMELD, uses `r length(cfs)` predictors.
Half of them, INR, (direct) bilirubin, and cystatin C are identical
or similar to the labortory measurements utilized in the classical MELD.
The direct bilirubin selected by our approach and the total bilirubin in MELD
are highly correlated (spearman correlation:
`r round(cor(amelddata$x, method = "spearman")["BILI_S", "BILID_S"])`) and
therefore interchangeable.

@@

In conclusion we developed a new model to predict the 90-day survival of
patients during the evaluation process for liver transplantation.
Our model, AMELD, extends the classical MELD predictors, INR,
(direct) bilirubin, cystatin C (instead of creatinine) by
synthesis parameters total protein and cholinesterase and
most importantly the inflammatory marker IL-6.
Using these `r length(cfs)` laboratory measurments the AMELD
outperforms the established MELD-Na and the recently published
MELD-Plus7 risk scores.
For a wider adoption a prospective multi center study for model
recalibration and validation is needed.
